---
title: "R Notebook"
output: html_notebook
---
 
Setup Environment 
```{r}
install.packages("randomForest")
install.packages("repr")
install.packages("caret")
install.packages("Metrics")
install.packages("AUC")
install.packages("cowplot")
suppressMessages(library(tidyverse))
suppressMessages(library(repr))
suppressMessages(library(randomForest))
suppressMessages(library(caret))
suppressMessages(library(cowplot))
suppressMessages(library(Metrics))
suppressMessages(library(AUC))

set.seed(123)
options(repr.plot.width=7, repr.plot.height=3)

```
Insert Data set
```{r}
library(haven)
ffqtot <- read_csv("C:/Users/copelaa/Downloads/ffqtot.csv")
View(ffqtot)
```
Sub-setting the data set to only the variables of interest
```{r}
AnalyticDataset = 
    ffqtot %>% select(
        currsm, bmi25,
        fkcal, atotfat, asat, aoleic, alinolc, achol, acarb, aprot, afiber, adca, adiron, adphos, adk, adna, advaiu, advare, adthiam, adribo, adniac, advc, atca, atiron, atvaiu, atthiri, atvc, asuptoc)
```

T-test between variables and BMI25 variable
```{r}

for (i in c("fkcal", "atotfat", "asat", "aoleic", "alinolc", "achol", "acarb", "aprot", "afiber", "adca", "adiron", "adphos", "adk", "adna", "advaiu", "advare", "adthiam", "adribo", "adniac", "advc", "atca", "atiron", "atvaiu", "atthiri", "atvc", "asuptoc", "currsm") ) {
    print(i)
}
```

```{r}
for (i in c("fkcal", "atotfat", "asat", "aoleic", "alinolc", "achol", "acarb", "aprot", "afiber", "adca", "adiron", "adphos", "adk", "adna", "advaiu", "advare", "adthiam", "adribo", "adniac", "advc", "atca", "atiron", "atvaiu", "atthiri", "atvc", "asuptoc", "currsm") ) {
    print(paste0(i, " ~ bmi25"))
}
```
Extract Pvalues from TTest
```{r}
ExtractedPValues = list()
for (i in c("fkcal", "atotfat", "asat", "aoleic", "alinolc", "achol", "acarb", "aprot", "afiber", "adca", "adiron", "adphos", "adk", "adna", "advaiu", "advare", "adthiam", "adribo", "adniac", "advc", "atca", "atiron", "atvaiu", "atthiri", "atvc", "asuptoc", "currsm") ) {
    ExtractedPValues[[i]] = t.test(formula = as.formula(paste0(i, " ~ bmi25")), data = AnalyticDataset)$p.value
}
ExtractedPValues = unlist(ExtractedPValues)
ExtractedPValues

```
Put Pvalues in Dataframe
```{r}
OutputTable = data.frame(
    X = names(ExtractedPValues), 
    Y = "bmi25", 
    PValue = ExtractedPValues
)
OutputTable<- OutputTable[order(ExtractedPValues),]
OutputTable

OutputStatSig<- subset.data.frame(OutputTable, PValue <= 0.05)
OutputStatSig
```
Looking at the Output table and the p values we can see that 
fkcal, aprot, asat, atotfat, aoleic, aoleic, atthiri are some of the values that have the lowest pvalues wiht a total of 12 variables that are significantly significant for predicting BMI25

Let's control the familywise error rate with the Bonferroni correction. 
- find the critical value (alpha) for an individual test by dividing the familywise error rate (0.05) by the number of tests. 
```{r}
 OutputTable$BonfP = 
        p.adjust(OutputTable$PValue, method = "bonferroni")

OutputStatSig<- subset.data.frame(OutputTable, BonfP <= (0.05/27))


```

Now after correcting we have 2 statistically significant variables 
Let's look at fkcal more...
```{r}
print(paste0("Sample mean:", round(mean(ffqtot$fkcal, na.rm=TRUE))))
print(paste0("Sample median:", median(ffqtot$fkcal, na.rm=TRUE)))
print(paste0("Sample variance:", var(ffqtot$fkcal,na.rm=TRUE)))
print(paste0("Sample Standard Deviation:", round(sd(ffqtot$fkcal,na.rm=TRUE))))
```
The average calorie intake for the sample is 2419 kcal

Visualize data
```{r}
fkcal<- data.frame(na.omit(ffqtot$fkcal))
colnames(fkcal) <- "Calories"

ggplot(fkcal, aes(x=Calories))+
      geom_histogram(bins=20, fill="light green", alpha=20, na.rm = TRUE)+
      geom_vline(aes(xintercept=mean(ffqtot$fkcal, na.rm = TRUE)),
              color = "red", size=0.5)+
      geom_vline(aes(xintercept= median(ffqtot$fkcal, na.rm = TRUE)),
              color= "red", linetype="dashed", size=0.5)
```
```{r}
ggplot(fkcal, aes(x="", y= Calories))+
  geom_boxplot(fill="light green", alpha=20) +
  stat_summary(fun.y=mean, colour="darkred", geom="point",size=2)
```
Relationship between BMI (continuous) and fkcal
```{r}
ggplot(ffqtot, aes(x=BMIBS, y=fkcal)) + 
    geom_point() +
    geom_smooth(method=lm, se=FALSE) +
    labs(x = "BMI",
         y = "Calories",
         title = "Relation between two variables")
```
Covariance and correlation between Calories and BMI (continuous variable not categorical)
```{r}
BMI.Cal<- na.omit(data.frame(ffqtot$BMIBS,ffqtot$fkcal))
BMI.Cal<- rename(BMI.Cal, "BMI" = "ffqtot.BMIBS")
BMI.Cal<- rename(BMI.Cal, "Calories" = "ffqtot.fkcal")
BMI_v<- as.vector(BMI.Cal$BMI)
Cal_v<- as.vector(BMI.Cal$Calories)
print(paste0("Covariance between BMI and Calorie Intake: ", round(cov(BMI_v, Cal_v),2)))
print(paste0("Correlation between BMI and Calorie Intake : ", round(cor(BMI_v,Cal_v),2)))
```
We can assume there is a relationship between BMI and Calorie intake Now why don't we look more at BMI as a categorical variable including aprot( ) heart disease and cholesterol 


Regression
```{r}
linearModel <- lm(BMIBS ~ fkcal + aprot, data=ffqtot)  # build linear regression model on full data
print(linearModel)

summary(linearModel)
```
Store results
```{r}
modelSummary <- summary(linearModel)  # capture model summary as an object
modelCoeffs <- modelSummary$coefficients  # model coefficients
beta.estimate <- modelCoeffs["afiber", "achol", "Estimate","Estimate"]  # get beta estimate for variables
std.error <- modelCoeffs["afiber","achol", "Std. Error"]  # get std.error for variables
t_value <- beta.estimate/std.error  # calc t statistic
p_value <- 2*pt(-abs(t_value), df=nrow(ffqtot)-ncol(ffqtot))  # calc p Value

```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
