---
title: "R Notebook"
output: html_notebook
---
 
Setup Enviroment 
```{r}
install.packages("randomForest")
install.packages("repr")
install.packages("caret")
install.packages("Metrics")
install.packages("AUC")
install.packages("cowplot")
suppressMessages(library(tidyverse))
suppressMessages(library(repr))
suppressMessages(library(randomForest))
suppressMessages(library(caret))
suppressMessages(library(cowplot))
suppressMessages(library(Metrics))
suppressMessages(library(AUC))

set.seed(123)
options(repr.plot.width=7, repr.plot.height=3)

```
Insert Data set
```{r}
library(haven)
fhs <- read_dta("C:/Users/copelaa/Downloads/fhs.dta")
View(fhs)
```
General data exploration and omitting Missing for general observations
```{r}
print(paste0("Sample mean:", round(mean(fhs$totchol3, na.rm=TRUE))))
print(paste0("Sample median:", median(fhs$totchol3, na.rm=TRUE)))
print(paste0("Sample variance:", var(fhs$totchol3,na.rm=TRUE)))
print(paste0("Sample Standard Deviation:", round(sd(fhs$totchol3,na.rm=TRUE))))
```
The average cholesterol level for the sample is 237 mg/dL

Visualize data
```{r}
totchol3<- data.frame(na.omit(fhs$totchol3))
colnames(totchol3) <- "Cholesterol"

ggplot(totchol3, aes(x=Cholesterol))+
      geom_histogram(bins=20, fill="light green", alpha=20, na.rm = TRUE)+
      geom_vline(aes(xintercept=mean(fhs$totchol3, na.rm = TRUE)),
              color = "red", size=0.5)+
      geom_vline(aes(xintercept= median(fhs$totchol3, na.rm = TRUE)),
              color= "red", linetype="dashed", size=0.5)
```
```{r}
ggplot(totchol3, aes(x="", y= Cholesterol))+
  geom_boxplot(fill="light green", alpha=20) +
  stat_summary(fun.y=mean, colour="darkred", geom="point",size=2)
```
Relationship between Cholesterol and Sex
```{r}
ggplot(fhs, aes(x=age3, y=totchol3)) + 
    geom_point() +
    geom_smooth(method=lm, se=FALSE) +
    labs(x = "Age",
         y = "Cholesterol",
         title = "Relation between two variables")
```
Covariance and correlation between Cholesterol and Age
```{r}
age.chol<- na.omit(data.frame(fhs$age3,fhs$totchol3))
age.chol<- rename(age.chol, "Age3" = "fhs.age3")
age.chol<- rename(age.chol, "Cholesterol" = "fhs.totchol3")
Age3_v<- as.vector(age.chol$Age3)
Chol_v<- as.vector(age.chol$Cholesterol)
print(paste0("Covariance between Cholesterol and Age: ", round(cov(Age3_v, Chol_v),2)))
print(paste0("Correlation between Cholesterol and Age: ", round(cor(Age3_v,Chol_v),2)))
```
We can assume there is not a relationship between age and Cholesterol. Now why don;t we look at Ischemic heart disease and cholesterol 

```{r}
ggplot(fhs, aes(x=stroke, y=death)) + 
    geom_point() +
    geom_smooth(method=lm, se=FALSE) +
    labs(x = "Stroke",
         y = "Death",
         title = "Relation between two variables")
```
```{r}
stroke.death<- na.omit(data.frame(fhs$stroke,fhs$death))
stroke.death<- rename(stroke.death, "Stroke" = "fhs.stroke")
stroke.death<- rename(stroke.death, "death" = "fhs.death")
Stroke_v<- as.vector(stroke.death$Stroke)
death_v<- as.vector(stroke.death$death)
print(paste0("Covariance between Death and stroke: ", round(cov(death_v,Stroke_v),2)))
print(paste0("Correlation between Death and stroke: ", round(cor(death_v,Stroke_v),2)))
```

Regression
1st we want to indicate that death is a dichotimous outcome 
```{r}
xtabs(~death + stroke, data = fhs)

logitmodel <- glm(death ~ stroke, data=fhs, family = binomial)  # build logistic regression model on full data
print(logitmodel)

summary(logitmodel)
```
Store results
```{r}
modelSummary <- summary(logitmodel)  # capture model summary as an object
modelCoeffs <- modelSummary$coefficients  # model coefficients
beta.estimate <- modelCoeffs["stroke", "Estimate"]  # get beta estimate for stroke
std.error <- modelCoeffs["stroke", "Std. Error"]  # get std.error for stroke
t_value <- beta.estimate/std.error  # calc t statistic
p_value <- 2*pt(-abs(t_value), df=nrow(fhs)-ncol(fhs))  # calc p Value

```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
