---
title: "Stand Your Ground Law and it's effect on Homicide rates in the US - Interrupted timeseries"
output: html_notebook
---
Stand Your Ground Law and it's effect on Homicide rates in the US - Interrupted timeseries 

1.  Bring in MERS data sets dowloaded from: http://wonder.cdc.gov/
    Centers for Disease Control and Prevention, National Center for Health Statistics. Underlying Cause of Death
    1999-2019 on CDC WONDER Online Database, released in 2020. Data are from the Multiple Cause of Death Files,        1999-2019, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics     Cooperative Program. Accessed at http://wonder.cdc.gov/ucd-icd10.html on Jun 12, 2021 9:26:53 PM
    
Install packages and open libraries as needed
```{r Install packages and libraries needed}
# install.packages("tidyverse")
# install.packages("RCurl") # Pull data from Github Raw version 
# install.packages("sensitivity")
# install.packages("ggplot2")
# install.packages("readxl")
# install.packages("reshape2")
library(readxl)
library(data.table)
library(ggplot2)
library(plotly)
library(ggplot2)
library (RCurl)
library (reshape2)
library(dplyr)
library(plotly)
library(rjson)
```

Insert Data. Most can be used as a guide on what to use for different scenarios. Importing text files (https://github.com/copelaa/myrepo/tree/master/Time_Series/Data.

Text files hosted on github and can be manually downloaded and opened from local machine, or pulled directly from Github. 
```{r}
##Text file of homicides in all states from 1999-2008
HOMICIDE_1999 = "https://raw.githubusercontent.com/copelaa/myrepo/master/Time_Series/Data/Underlying%20Cause%20of%20Death%2C%201999-2008.txt"

HOMICIDE_1999 = read.delim(HOMICIDE_1999, header = TRUE)

##Text file of homicides in all states from 2009-2019
HOMICIDE_2009 = "https://raw.githubusercontent.com/copelaa/myrepo/master/Time_Series/Data/Underlying%20Cause%20of%20Death%2C%202009-2019.txt"

##Country file to plot over country image

data <- fromJSON(file="https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json")
data$features[[1]]

HOMICIDE_2009 = read.delim(HOMICIDE_2009, header = TRUE)

#Merge dataframes so years are 1999->2019
total_HOMICIDE <- rbind(HOMICIDE_1999, HOMICIDE_2009)
```

Add Stand Your Ground Starting dates for States:
```{r}
total_HOMICIDE$SYG_Start <- NA

#Change Date format to show as dates in Dataframe
total_HOMICIDE$SYG_Start<-as.Date(total_HOMICIDE$SYG_Start,origin = "1970-01-01", format = "%m/%d/%y")

#Alabama
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Alabama"] <- as.Date("06/01/2006",format = "%m/%d/%y")

#Alaska
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Alaska"] <- as.Date("09/13/2006",format = "%m/%d/%y")

#Arizona
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Arizona"] <- as.Date("04/24/2006",format = "%m/%d/%y")

#Arkansas
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Arkansas"] <- as.Date("03/03/2021",format = "%m/%d/%y")

#Colorado
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Colorado"] <- as.Date("01/01/2019",format = "%m/%d/%y")

#Florida
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Florida"] <- as.Date("10/01/2005",format = "%m/%d/%y")

#Georgia
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Georgia"] <- as.Date("07/01/2006",format = "%m/%d/%y")

#Idaho
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Idaho"] <- as.Date("07/01/2006",format = "%m/%d/%y")

#Illinois
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Illinois"] <- as.Date("07/28/2004",format = "%m/%d/%y")

#Indiana 
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Indiana"] <- as.Date("07/01/2006",format = "%m/%d/%y")

#Iowa
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Iowa"] <- as.Date("04/18/2017",format = "%m/%d/%y")

#Kansas
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Kansas"] <- as.Date("07/01/2006",format = "%m/%d/%y")

#Kentucky
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Kentucky"] <- as.Date("07/12/2006",format = "%m/%d/%y")

#Louisiana
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Louisiana"] <- as.Date("08/15/2006",format = "%m/%d/%y")

#Michigan
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Michigan"] <- as.Date("10/01/2006",format = "%m/%d/%y")

#Mississippi
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Mississippi"] <- as.Date("07/01/2006",format = "%m/%d/%y")

#Missouri
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Missouri"] <- as.Date("08/28/2007",format = "%m/%d/%y")

#Montana
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Montana"] <- as.Date("04/27/2009",format = "%m/%d/%y")

#Nevada
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Nevada"] <- as.Date("10/01/2011",format = "%m/%d/%y")

#New Hampshire 
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "New Hampshire"] <- as.Date("11/13/2011",format = "%m/%d/%y")

#North Carolina
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "North Carolina"] <- as.Date("12/01/2011",format = "%m/%d/%y")

#North Dakota
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "North Dakota"] <- as.Date("08/01/2007",format = "%m/%d/%y")

#Ohio
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Ohio"] <- as.Date("09/09/2008",format = "%m/%d/%y")

#Oklahoma
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Oklahoma"] <- as.Date("11/01/2006",format = "%m/%d/%y")

#Pennsylvania
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Pennsylvania"] <- as.Date("06/01/2011",format = "%m/%d/%y")

#South Carolina
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "South Carolina"] <- as.Date("06/09/2006",format = "%m/%d/%y")

#South Dakota
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "South Dakota"] <- as.Date("07/01/2006",format = "%m/%d/%y")

#Tennessee
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Tennessee"] <- as.Date("05/22/2007",format = "%m/%d/%y")

#Texas
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Texas"] <- as.Date("09/01/2007",format = "%m/%d/%y")

#Utah
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Utah"] <- as.Date("03/02/1994",format = "%m/%d/%y")

#West Virginia
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "West Virginia" ]<- as.Date("02/28/2008",format = "%m/%d/%y")
 
#Wyoming                        
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Wyoming"] <- as.Date("07/01/2008",format = "%m/%d/%y")     

#Check States with No SYG Laws
NO_SYG<- data.frame(total_HOMICIDE$State[is.na(total_HOMICIDE$SYG_Start)])

colnames(NO_SYG)[1] <- "State"

NO_SYG<- NO_SYG[!duplicated(NO_SYG[ ,c("State")]),]
  

#Check States with SYG Laws
SYG<- data.frame(total_HOMICIDE[ ,c(4,23)])
SYG<- na.omit(SYG)
SYG<- as.data.frame(SYG[ ,1])
colnames(SYG)[1] <- "State"
SYG<- SYG[!duplicated(SYG[ ,c("State")]),]       
States<- total_HOMICIDE[!duplicated(total_HOMICIDE[ ,c("State")]),4]
```

Data Exploration: 

Plot Homicide level over USA Image
```{r}
test <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
test$hover <- with(test, paste(state, '<br>', "Beef", beef, "Dairy", dairy, "<br>",
                           "Fruits", total.fruits, "Veggies", total.veggies,
                           "<br>", "Wheat", wheat, "Corn", corn))
HOMICIDE_2010 <- as.data.frame(total_HOMICIDE[total_HOMICIDE$Year== 2010, ])
HOMICIDE_2010$Deaths<- na.omit(HOMICIDE_2010$Deaths)
colnames(HOMICIDE_2010)[12]<- "Deaths"
HOMICIDE_2010<- subset.data.frame(HOMICIDE_2010[c("Year", "State", "Deaths", "Population")])
HOMICIDE_2010<- HOMICIDE_2010$State%>%
  summarize(Deaths=sum(Deaths))

df_new = df.groupby(df['id']).aggregate(aggregation_functions)
HOMICIDE_2010$hover <- with(HOMICIDE_2010, paste(State, '<br>', "Beef", beef, "Dairy", dairy, "<br>",
                           "Fruits", total.fruits, "Veggies", total.veggies,
                           "<br>", "Wheat", wheat, "Corn", corn))
# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

fig <- plot_geo(df, locationmode = 'USA-states')
fig <- fig %>% add_trace(
    z = ~total.exports, text = ~hover, locations = ~code,
    color = ~total.exports, colors = 'Purples'
  )
fig <- fig %>% colorbar(title = "Homicide by firearm")
fig <- fig %>% layout(
    title = '1999-2019 Homicide Rates in the USA <br>(Hover for breakdown)',
    geo = g
  )

fig
```

