---
title: "Stand Your Ground Law and it's effect on Homicide rates in the US - Interrupted timeseries"
output: html_notebook
---
Stand Your Ground Law and it's effect on Homicide rates in the US - Interrupted timeseries 

1.  Bring in MERS data sets downloaded from: http://wonder.cdc.gov/
    Centers for Disease Control and Prevention, National Center for Health Statistics. Underlying Cause of Death
    1999-2019 on CDC WONDER Online Database, released in 2020. Data are from the Multiple Cause of Death Files,        1999-2019, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics     Cooperative Program. Accessed at http://wonder.cdc.gov/ucd-icd10.html on Jun 12, 2021 9:26:53 PM
    
Install packages and open libraries as needed
```{r Install packages and libraries needed}
# install.packages("tidyverse")
# install.packages("RCurl") # Pull data from Github Raw version 
# install.packages("sensitivity")
# install.packages("ggplot2")
# install.packages("readxl")
# install.packages("reshape2")
# install.packages("lubridate")
# install.packages("pander")
# install.packages("stargazer")
# install.packages("scales")
# install.packages("olsrr")
library(olsrr)
library(scales)
library(stargazer)
library(pander)
library(lubridate)
library(readxl)
library(data.table)
library(ggplot2)
library(plotly)
library(ggplot2)
library (RCurl)
library (reshape2)
library(dplyr)
library(plotly)
library(rjson)
library(tidyr)
library(reshape2)
```

Insert Data. Most can be used as a guide on what to use for different scenarios. Importing text files (https://github.com/copelaa/myrepo/tree/master/Time_Series/Data.

Text files hosted on github and can be manually downloaded and opened from local machine, or pulled directly from Github. 
```{r}
##Text file of homicides in all states by county and race from 1999-2008
HOMICIDE_1999 = "https://raw.githubusercontent.com/copelaa/myrepo/master/Time_Series/Data/Underlying%20Cause%20of%20Death%2C%201999-2008.txt"

HOMICIDE_1999 = read.delim(HOMICIDE_1999, header = TRUE)

##Text file of homicides in all states by county and race from 2009-2019
HOMICIDE_2009 = "https://raw.githubusercontent.com/copelaa/myrepo/master/Time_Series/Data/Underlying%20Cause%20of%20Death%2C%202009-2019.txt"
HOMICIDE_2009 = read.delim(HOMICIDE_2009, header = TRUE)

##Text file of homicides in all states 1999-2019
HOMICIDE_States = "https://raw.githubusercontent.com/copelaa/myrepo/master/Time_Series/Data/Underlying%20Cause%20of%20Death%2C%201999-2019%20States.txt"
HOMICIDE_States = read.delim(HOMICIDE_States, header = TRUE)

##Country file to plot over country image
data <- fromJSON(file="https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json")
data$features[[1]]


#Merge dataframes so years are 1999->2019
total_HOMICIDE <- rbind(HOMICIDE_1999, HOMICIDE_2009)
```

Add Stand Your Ground Starting dates for States:
```{r}
total_HOMICIDE$SYG_Start <- NA



#Alabama
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Alabama"] <- as.Date("06/01/2006",format = "%m/%d/%Y")

#Alaska
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Alaska"] <- as.Date("09/13/2006",format = "%m/%d/%Y")

#Arizona
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Arizona"] <- as.Date("04/24/2006",format = "%m/%d/%Y")

#Arkansas
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Arkansas"] <- as.Date("03/03/2021",format = "%m/%d/%Y")

#Colorado
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Colorado"] <- as.Date("01/01/2019",format = "%m/%d/%Y")

#Florida
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Florida"] <- as.Date("10/01/2005",format = "%m/%d/%Y")

#Georgia
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Georgia"] <- as.Date("07/01/2006",format = "%m/%d/%Y")

#Idaho
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Idaho"] <- as.Date("07/01/2006",format = "%m/%d/%Y")

#Illinois
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Illinois"] <- as.Date("07/28/2004",format = "%m/%d/%Y")

#Indiana 
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Indiana"] <- as.Date("07/01/2006",format = "%m/%d/%Y")

#Iowa
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Iowa"] <- as.Date("04/18/2017",format = "%m/%d/%Y")

#Kansas
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Kansas"] <- as.Date("07/01/2006",format = "%m/%d/%Y")

#Kentucky
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Kentucky"] <- as.Date("07/12/2006",format = "%m/%d/%Y")

#Louisiana
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Louisiana"] <- as.Date("08/15/2006",format = "%m/%d/%Y")

#Michigan
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Michigan"] <- as.Date("10/01/2006",format = "%m/%d/%Y")

#Mississippi
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Mississippi"] <- as.Date("07/01/2006",format = "%m/%d/%Y")

#Missouri
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Missouri"] <- as.Date("08/28/2007",format = "%m/%d/%Y")

#Montana
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Montana"] <- as.Date("04/27/2009",format = "%m/%d/%Y")

#Nevada
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Nevada"] <- as.Date("10/01/2011",format = "%m/%d/%Y")

#New Hampshire 
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "New Hampshire"] <- as.Date("11/13/2011",format = "%m/%d/%Y")

#North Carolina
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "North Carolina"] <- as.Date("12/01/2011",format = "%m/%d/%Y")

#North Dakota
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "North Dakota"] <- as.Date("08/01/2007",format = "%m/%d/%Y")

#Ohio
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Ohio"] <- as.Date("09/09/2008",format = "%m/%d/%Y")

#Oklahoma
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Oklahoma"] <- as.Date("11/01/2006",format = "%m/%d/%Y")

#Pennsylvania
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Pennsylvania"] <- as.Date("06/01/2011",format = "%m/%d/%Y")

#South Carolina
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "South Carolina"] <- as.Date("06/09/2006",format = "%m/%d/%Y")

#South Dakota
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "South Dakota"] <- as.Date("07/01/2006",format = "%m/%d/%Y")

#Tennessee
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Tennessee"] <- as.Date("05/22/2007",format = "%m/%d/%Y")

#Texas
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Texas"] <- as.Date("09/01/2007",format = "%m/%d/%Y")

#Utah
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Utah"] <- as.Date("03/02/1994",format = "%m/%d/%Y")

#West Virginia
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "West Virginia" ]<- as.Date("02/28/2008",format = "%m/%d/%Y")
 
#Wyoming                        
total_HOMICIDE$SYG_Start[total_HOMICIDE$State == "Wyoming"] <- as.Date("07/01/2008",format = "%m/%d/%Y")     

#Change Date format to show as dates in Dataframe
total_HOMICIDE$SYG_Start<- as_date(total_HOMICIDE$SYG_Start, origin = lubridate::origin)

```

While the data frame comes with a crude adjusted death rate for Death below is how to normalize death based on population
```{r Normalize}

# Calculate death rates per state directly standardized for population
total_HOMICIDE <- mutate(total_HOMICIDE, Crude.Rate.Death = 100000 * Deaths / Population)

HOMICIDE_States <- mutate(HOMICIDE_States, Crude.Rate.Death = 100000 * Deaths / Population)
```


```{r Create Dataframes for SYG and Non-SYG States}

#Check States with No SYG Laws
NO_SYG<- data.frame(total_HOMICIDE$State[is.na(total_HOMICIDE$SYG_Start)])
colnames(NO_SYG)[1] <- "State"
NO_SYG<- NO_SYG[!duplicated(NO_SYG[ ,c("State")]),]
  

#Check States with SYG Laws
SYG<- data.frame(total_HOMICIDE[ ,c(4,22)])
SYG<- na.omit(SYG)
SYG<- as.data.frame(SYG[ ,1])
colnames(SYG)[1] <- "State"
SYG<- SYG[!duplicated(SYG[ ,c("State")]),]       
States<- total_HOMICIDE[!duplicated(total_HOMICIDE[ ,c("State")]),4]

#Separate year from datetime:
total_HOMICIDE$SYG <- format(total_HOMICIDE$SYG_Start, format="%Y")

```

State Dataframe for actual use for TS analysis 
```{r Divide State data by States that have SYG laws and do not}
HOMICIDE_States$SYG_Start <- NA

#Add SYG Year and SYG Start date to State Dataframe from Total Homicide dataframe
SYG_Date = data.frame(State=c(total_HOMICIDE$State) , SYG= 
                        as.numeric(total_HOMICIDE$SYG), 
                      SYG_Start=as.Date(total_HOMICIDE$SYG_Start))
SYG_Date <- SYG_Date[!duplicated(SYG_Date[ ,c("State")]),1:3]
row.names(SYG_Date) <- SYG_Date$State                     # Convert column to row names
SYG_Date$State<-NULL
HOMICIDE_States$SYG<-SYG_Date[HOMICIDE_States$State,"SYG"]
HOMICIDE_States$SYG_Start<-SYG_Date[HOMICIDE_States$State, "SYG_Start"]

#Check States with No SYG Laws
NO_SYG_states<- data.frame(HOMICIDE_States$State[is.na(HOMICIDE_States$SYG_Start)])
colnames(NO_SYG_states)[1] <- "State"
NO_SYG_states<- NO_SYG_states[!duplicated(NO_SYG_states[ ,c("State")]),]
  
#Check States with SYG Laws
SYG_states<- data.frame(HOMICIDE_States[ ,c(4,18)])
SYG_states<- na.omit(SYG_states)
SYG_states<- as.data.frame(SYG_states[ ,1])
colnames(SYG_states)[1] <- "State"
SYG_states<- SYG_states[!duplicated(SYG_states[ ,c("State")]),]       
States<- HOMICIDE_States[!duplicated(HOMICIDE_States[ ,c("State")]),4]

#Make an Intervention column
HOMICIDE_States$I<- NA

HOMICIDE_States$I<- ifelse(HOMICIDE_States$Year >= HOMICIDE_States$SYG & !is.na(HOMICIDE_States$SYG),1,0)

#Add End date
HOMICIDE_States$END<- c(2019)

#Create Time Since intervention 
HOMICIDE_States$P <- (HOMICIDE_States$END-HOMICIDE_States$SYG)

#Drop rows 974 on. (all disclaimer information)
HOMICIDE_States<-HOMICIDE_States[-c(974:1046),] 
```

Data Exploration: 

```{r Plot Homicide level over USA Image}

test <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")

HOMICIDE_States_2009<- HOMICIDE_States[ which(HOMICIDE_States$Year=="2009"), ]
# ADD State codes to data frame 2009
x=data.frame(row.names=c(test$state) , val=c(test$code))
HOMICIDE_States_2009$State_AB<- x[HOMICIDE_States_2009$State,]

#Map on US Map
HOMICIDE_States_2009$hover <- with(HOMICIDE_States_2009, paste(State, '<br>', "Crude.Rate.Death", Crude.Rate.Death, "Population", Population, "<br>",
                           "Age Adjusted Rate", Age.Adjusted.Rate))
# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

fig <- plot_geo(HOMICIDE_States_2009, locationmode = 'USA-states')
fig <- fig %>% add_trace(
    z = ~Crude.Rate.Death, text = ~hover, locations = ~State_AB,
    color = ~Crude.Rate.Death, colors = 'Purples'
  )
fig <- fig %>% colorbar(title = "Homicide by firearm")
fig <- fig %>% layout(
    title = '2009 Homicide Rates in the USA <br>(Hover for breakdown)',
    geo = g
  )

fig

```



```{r Plot intial time series}

#Plot SYG States and deaths
HOMICIDE_STATES_SYG <- HOMICIDE_States %>% filter(!is.na(SYG))
HOMICIDE_STATES_NOSYG<- HOMICIDE_States %>% filter(is.na(SYG))

plot(HOMICIDE_STATES_SYG$Year, HOMICIDE_STATES_SYG$Crude.Rate.Death,
      bty="n", pch=19, col="gray",
      ylim = c(min(HOMICIDE_STATES_SYG$Deaths, na.rm = TRUE), max(HOMICIDE_STATES_SYG$Crude.Rate.Death, na.rm = TRUE)), xlim=c(1999,2020),
      xlab = "Time (years)", 
      ylab = "Homicide Deaths by Firearms")

# Line marking the interruption
abline( v=2006, col="firebrick", lty=2 )
text( 200, 300, "Start of Stand your Ground Laws", col="firebrick", cex=1.3, pos=4 )

# Add the regression line
ts <- lm(Crude.Rate.Death ~ Year + I + P, data=HOMICIDE_STATES_SYG)
lines(HOMICIDE_STATES_SYG$Year, ts$fitted.values, col="steelblue", lwd=2)
summary(ts)


#Plot no SYG Law States 
plot(HOMICIDE_STATES_NOSYG$Year, HOMICIDE_STATES_NOSYG$Crude.Rate.Death,
      bty="n", pch=19, col="gray",
      ylim = c(min(HOMICIDE_STATES_NOSYG$Deaths, na.rm = TRUE), max(HOMICIDE_STATES_NOSYG$Crude.Rate.Death, na.rm = TRUE)), xlim=c(1999,2020),
      xlab = "Time (years)", 
      ylab = "Homicide Deaths by Firearms")

# Line marking the interruption
abline( v=2006, col="firebrick", lty=2 )
text( 200, 300, "Start of Stand your Ground Laws", col="firebrick", cex=1.3, pos=4 )

# Add the regression line
ts_NO <- lm(Crude.Rate.Death ~ Year, data=HOMICIDE_STATES_NOSYG)
lines(HOMICIDE_STATES_NOSYG$Year, ts_NO$fitted.values, col="steelblue", lwd=2)
summary(ts_NO)

```


Column	Variable name	Description
Deaths: Deaths by firearms, not including Suicide
T: Year-Time (from 1999 to 2019)
I: Stand Your Ground Intervention.	Observation post (=1) and pre (=0) intervention
P: Time Since Intervention Time passed since the intervention in years
Our model is based on the equation (1.1):

Y=b0+b1∗Time+b2∗Intervention+b3∗TimeSinceIntervention+e

```{r Timeseries}
regTS = lm (Crude.Rate.Death ~ Year + I + P, data=HOMICIDE_STATES_SYG)# Our time series model

suppressWarnings(stargazer(regTS,type="text"))

```

Lets look at a specific states as the SYG laws went into effect at multiple times
```{r Ohio}

#Ohio 
HOMICIDE_Ohio_SYG<- HOMICIDE_STATES_SYG[which(HOMICIDE_STATES_SYG$State=="Ohio"), ]

#Plot SYG States and deaths
plot(HOMICIDE_Ohio_SYG$Year, HOMICIDE_Ohio_SYG$Crude.Rate.Death,
      bty="n", pch=19, col="gray",
      ylim = c(min(HOMICIDE_Ohio_SYG$Crude.Rate.Death, na.rm = TRUE), max(HOMICIDE_Ohio_SYG$Crude.Rate.Death, na.rm = TRUE)), xlim=c(1999,2020),
      xlab = "Time (years)", 
      ylab = "Homicide Deaths by Firearms")

# Line marking the interruption
abline( v=2008, col="firebrick", lty=2 )
text( 200, 300, "Start of Stand your Ground Laws", col="firebrick", cex=1.3, pos=4 )

# Add the regression line
ts_Ohio <- lm(Crude.Rate.Death ~ Year + I + P, data=HOMICIDE_Ohio_SYG)
lines(HOMICIDE_Ohio_SYG$Year, ts_Ohio$fitted.values, col="steelblue", lwd=2)
summary(ts_Ohio)
```

```{r Texas}
#Texas 
HOMICIDE_Texas_SYG<- HOMICIDE_STATES_SYG[which(HOMICIDE_STATES_SYG$State=="Texas"), ]

#Plot SYG States and deaths
plot(HOMICIDE_Texas_SYG$Year, HOMICIDE_Texas_SYG$Crude.Rate.Death,
      bty="n", pch=19, col="gray",
      ylim = c(min(HOMICIDE_Texas_SYG$Crude.Rate.Death, na.rm = TRUE), max(HOMICIDE_Texas_SYG$Crude.Rate.Death, na.rm = TRUE)), xlim=c(1999,2020),
      xlab = "Time (years)", 
      ylab = "Homicide Deaths by Firearms")

# Line marking the interruption
abline( v=2007, col="firebrick", lty=2 )
text( 200, 300, "Start of Stand your Ground Laws", col="firebrick", cex=1.3, pos=4 )

# Add the regression line
ts_Texas <- lm(Crude.Rate.Death ~ Year + I + P, data=HOMICIDE_Texas_SYG)
lines(HOMICIDE_Texas_SYG$Year, ts_Texas$fitted.values, col="steelblue", lwd=2)
summary(ts_Texas)

regTS_Texas = lm (Crude.Rate.Death ~ Year + I + P, data=HOMICIDE_Texas_SYG)# Our time series model

suppressWarnings(stargazer(regTS_Texas,type="text"))

```
Now, we can calculate the outcome 8 years after the intervention. Time will be equal to 2015 while Time Since Treatment will be equal to 8. Treatment is always equal to 1 since it is a dummy variable indicating pre or post intervention.

Y=b0+b1∗2015+b2∗1+b3∗8+e(2.3)

We can add the new point at t = 2015 to the graph:
```{r Texas further exploration}
Texas2<- as.data.frame( cbind( Year = 2015, I = 1, P = 8 )) # New data

y2 <- predict( regTS_Texas, Texas2 ) # We predict the new outcome

# We plot our initial observations, the column Y in our dataset
plot( HOMICIDE_Texas_SYG$Year, HOMICIDE_Texas_SYG$Crude.Rate.Death,
      bty="n",
      col = gray(0.5,0.5), pch=19,
      xlim = c(1999,2020), 
      ylim = c(0,10),
      xlab = "Time(Years)", 
      ylab = "Deaths")

# We add a point showing the level of Death at time = 2015
points(2015, y2, col = "dodgerblue4", pch = 19, bg = "red", cex = 2)

# Label for our predicted outcome
text(2015, y2, labels = "Year = 2015", pos = 4, cex = 1)

# Line marking the interruption
abline( v=2008, col="red", lty=2 )
```

Calculate the counterfactual for any point in time. The counterfactual at Time = 2015 is the level of Deaths at that point in time if the intervention had not occurred.

Note the changes in our equation (2.3): Treatment and Time Since Treatment are both equal to 0.

Y=b0+b1∗230+b2∗0+b3∗0+e

As before, we plot the results on a graph. Y is our predict value and C is the counterfactual.

```{r Texas Counterfactual}
Texas3 <- as.data.frame(cbind( Year= 2015, I = 0, P = 0)) # Data if the intervention does not occur

y3 <- predict(regTS_Texas, Texas3) #Counterfactual

# We plot our initial observations, the column Y in our dataset
plot( HOMICIDE_Texas_SYG$Year, HOMICIDE_Texas_SYG$Crude.Rate.Death,
      bty="n",
      col = gray(0.5,0.5), pch=19,
      xlim = c(1999, 2020), 
      ylim = c(0, 6),
      xlab = "Time (Years)", 
      ylab = "Death")

# We add a  point showing the level of Death at time = 2015
points(2015, y2, col = "dodgerblue4", pch = 15, bg = "red", cex = 2)

# We add a point indicating the counterfactual
points(2015, y3, col = "darkorange2", pch = 15, cex = 2)

# Label for our predicted outcome
text(2015, y2, labels = "Deaths = 1100 at t = 2015", pos = 2, cex = 1)

#Label for the counterfactual 
text(2015, y3, labels = "C.Deaths = 1268 at t = 2015", pos = 2, cex = 1)

# Line marking the interruption
abline( v=2008, col="red", lty=2 )
```
plotting all predicted outcomes and their counterfactuals. We can see that there is an increase in the number of deaths after the intervention. The dashed line represents the counterfactual.
```{r Texas Plot Inturrupted time series}
pred1_Texas <- predict(regTS_Texas, HOMICIDE_Texas_SYG) 
# To estimate all predicted values of Death, we just use our dataset
 
datanew_Texas<- as.data.frame(cbind(Year = rep(1999 : 2020), I = rep(0), P = rep(0))) 
# Create a new dataset where Treatment and Time Since Treatment are equal to 0 as the intervention did not occur.

pred2_Texas <- predict(regTS_Texas, datanew_Texas)

# Predict the counterfactuals
plot( HOMICIDE_Texas_SYG$Year, HOMICIDE_Texas_SYG$Crude.Rate.Death,
      bty="n",
      col = gray(0.5,0.5), pch=19,
      xlim = c(1999, 2020), 
      ylim = c(0, 6),
      xlab = "Time (Years)", 
      ylab = "Deaths")

lines( rep(1999:2008), pred1_Texas[1:10], col="dodgerblue4", lwd = 3 )
lines( rep(2008:2020), pred1_Texas[10:22], col="dodgerblue4", lwd = 3 )
lines( rep(2008:2020), pred2_Texas[10:22], col="darkorange2", lwd = 3, lty = 5 ) 

text(2000, 800, labels = "Predicted values", pos = 4, cex = 1, col = "dodgerblue3")
text(2015, 1400, labels = "Counterfactual", pos = 4, cex = 1, col = "darkorange2")

# Line marking the interruption
abline( v=2008, col="darkorange2", lty=2 )

```
Note that the coefficients obtained from equation (2.1) do not tell us if the difference between each point (the predicted outcomes) and its counterfactual is statistically significant. It only tells you if there is an immediate change after the intervention and the slope has changed after the intervention.

There is no statistical test to look at whether there is a statistically significant difference between a predicted outcome and its counterfactual. This has important implications. It could be that effect of the intervention varies over time. In this case, there is a drop in homicide deaths a few years after the intervention. For instance, it could be that the law was doing as it intended to do right after it was made law. However we also have to consider outside factors that may contribute to the drop in homicides from 2008-2012.

NO SYG Group Compared to SYG Groups
```{r Get Dataframe ready for no SYG group}
NO_SYG_Var<- c("Year","State","Crude.Rate.Death","I","P")
NO_SYG<-HOMICIDE_STATES_NOSYG[NO_SYG_Var]
NO_SYG["P"]=0

plot(NO_SYG$Year, NO_SYG$Crude.Rate.Death,
      bty="n", pch=19, col="gray",
      ylim = c(min(NO_SYG$Crude.Rate.Death, na.rm = TRUE), max(NO_SYG$Crude.Rate.Death, na.rm = TRUE)), xlim=c(1999,2020),
      xlab = "Time (years)", 
      ylab = "Homicide Deaths by Firearms")

#find regression of No SYG states
ts_NOSYG <- lm(Crude.Rate.Death ~ Year, data=NO_SYG)
lines(NO_SYG$Year, ts_NOSYG$fitted.values, col="steelblue", lwd=2)
summary(ts_NOSYG)

regTS_NOSYG = lm(Crude.Rate.Death ~ Year, data=NO_SYG)# Our time series model

suppressWarnings(stargazer(regTS_NOSYG,type="text"))
```

```{r Reggresion for all states}
#Plot all States 
plot(HOMICIDE_States$Year, HOMICIDE_States$Crude.Rate.Death,
      bty="n", pch=19, col="gray",
      ylim = c(min(HOMICIDE_States$Crude.Rate.Death, na.rm = TRUE), max(HOMICIDE_States$Crude.Rate.Death, na.rm = TRUE)), xlim=c(1999,2020),
      xlab = "Time (years)", 
      ylab = "Homicide Deaths by Firearms")

# Line marking the interruption
abline( v=2006, col="firebrick", lty=2 )
text( 200, 300, "Start of Stand your Ground Laws", col="firebrick", cex=1.3, pos=4 )

# Add the regression line
lines(HOMICIDE_STATES_NOSYG$Year, ts_NOSYG$fitted.values, col="steelblue", lwd=2)
lines(HOMICIDE_STATES_SYG$Year, ts$fitted.values, col="firebrick", lwd=2)

summary(ts_NO)
```

 
```{r Check for correlation among points}

attach(HOMICIDE_States)
plot(Crude.Rate.Death, Year, main="Scatterplot Example",
   xlab="Deaths ", ylab="Year", pch=19)
regTS_OK <- lm(Crude.Rate.Death ~ Year, data = HOMICIDE_States)
# Calculate a simple regression where birth rate depends only on time 
print(regTS_OK)
summary(regTS_OK)
plot(resid( regTS_OK ))
ols_plot_cooksd_bar(regTS_OK)

```

Drop California to see if it makes a substantial difference in

