---
title: "Stand Your Ground Law and it's effect on Homicide rates in the US - Interrupted timeseries"
output: html_notebook
---
Stand Your Ground Law and it's effect on Homicide rates in the US - Interrupted timeseries 

1.  Bring in MERS data sets dowloaded from: http://wonder.cdc.gov/
    Centers for Disease Control and Prevention, National Center for Health Statistics. Underlying Cause of Death
    1999-2019 on CDC WONDER Online Database, released in 2020. Data are from the Multiple Cause of Death Files,        1999-2019, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics     Cooperative Program. Accessed at http://wonder.cdc.gov/ucd-icd10.html on Jun 12, 2021 9:26:53 PM
    
Install packages and open libraries as needed
```{r Install packages and libraries needed}
# install.packages("tidyverse")
# install.packages("RCurl") # Pull data from Github Raw version 
# install.packages("sensitivity")
# install.packages("ggplot2")
# install.packages("readxl")
# install.packages("reshape2")
library(readxl)
library(deSolve)
library(data.table)
library(data.table)
library(ODEsensitivity)# If you want to use the s
library(ggplot2)
library(plotly)
library(lattice)#Making heat maps
library(latticeExtra) #make Heat maps
library(ggplot2)
library (RCurl)
library (reshape2)
```

Insert Data. Most can be used as a guide on what to use for different scenarios. Importing text files (https://github.com/copelaa/myrepo/tree/master/Time_Series/Data.

Text files hosted on github and can be manually downloaded and opened from local machine, or pulled directly from Github. 
```{r}
##Text file of homicides in all states from 1999-2008
HOMICIDE_1999 = "https://github.com/copelaa/myrepo/blob/master/Time_Series/Data/Underlying%20Cause%20of%20Death%2C%202019-2008.txt"

HOMICIDE_1999 = read.delim("https://github.com/copelaa/myrepo/blob/master/Time_Series/Data/Underlying%20Cause%20of%20Death%2C%202019-2008.txt", header = TRUE)

##Text file of homicides in all states from 2009-2019
HOMICIDE_2009 = "https://github.com/copelaa/myrepo/blob/master/Time_Series/Data/Underlying%20Cause%20of%20Death%2C%202009-2019.txt"

HOMICIDE_2009 = read.delim(HOMICIDE_2009, header = TRUE)

```

